A Logic-Based Framework for Digital Signal Processing

### 2008-06-06 paper/paper.tex
draft version of the paper

This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency. This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one. There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog \cite{lambda} or Mercury \cite{mercury}. However, our approach results to be more suitable to develop an open platform for research and development. External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface. Furthermore, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.

@@

\begin{itemize}
 \item signal: this type represents an audio signal in a physical timeline. T
\end{itemize}

@@

These standards allow us to work with different modules without taking care of the specific underlying APIs types or implementation, so that we achieve a logic framework with dedicated semantics to perform audio analysis tasks as showed in \ref{fig:workflow}.

\begin{figure}
\centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
\caption{Workflow specification of the modules}
\label{fig:workflow}
\end{figure}

We can then execute the following task as a SWI-Prolog query. \textit{The input audio file is decoded and then we apply a pre-processing in which we amplify the signal. We then apply a key detection algorithm over the signal to retrieve tonal regions}

my_task(Input, ['Parameter'(gain, 2)], KeyProfile):-
	decode(InputFile, Signal),
	amplify(Signal, 'Parameter'(gain, 2), SignalX2),
	key(SignalX2, KeyProfile).

KeyProfile = ['Feature'(key, 'Timestamp'(0, 23.45), 'Amajor')|...].

\end{verbatim}

@@

\section{Towards a computation reasoner}\label{sec:reasoner}

So far we have presented how SWI-DSP provides a vocabulary of logic predicates to perform audio analysis computations, but this logic front-end is still inappropiate for such tasks. We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification. The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users. The task-oriented module, which is itself very layered, provides high-level predicates which are missing important options at setting-up.

In addition to this, we require a mechanism to update results and keep track of computation sequences, so that computations are not restricted to one SWI-Prolog session, allowing us to reuse and share them. This section tackles this problem by extending SWI-DSP as an expert system able to interpret data within defined ``DSP scenarios'' and by introducing Concurrent Transaction Logic on top of our extended predicate-logic language.

\subsection{DSP scenarios}\label{subsec:scenarios}

Our DSP workflow is strongly constrained by the predicate instantation patterns which clearly establishes a hierarchy of users depeding on the level of abstraction we are dealing with. Applications provide the user with a nice graphical or command line interface to control the computations. We need therefore to set a ``nice view'' of the system to the potential users which are not (and do not need to be) related to the SWI-DSP or the libraries implementation.

We address this problem by defining ``scenarios'' that control and delimit the user's view of the system without restricting significantly his options. Any computation over an audio signal can be seen as a function input/ouput with a specific setup. The ``scenario'' characterizes such a family of funtions in a generic way so the user just needs to configure the scenario and provide the input to capture the output. Such computations can be wrapped as a signal \textbf{transform} which is declared as a SWI-Prolog functor:

'Transform'(TransformType, 'Engine'(API, PluginName, Version), SampleRate, StepSize, BlockSize, ListOfParameters, Program).

\begin{itemize}

 \item TransformType indicates which sort of task we are carrying out (feature extraction, audio effect...)
 \item 'Engine'(A, Pn, V) identifies the computation engine in charge of the processing of the transform
 \item SampleRate indicates the input sample rate in case it is not constant
 \item StepSize and BlockSize are the framing parameters
 \item ListOfParameters to setup the engine
 \item Program is a predetermined configuration for the engine
\end{itemize}

Then each task module hardcodes the necessary knowledge in form of interpretation rules to read the tranform scenario and set the up the corresponding engine. The predicate \textit{transform(+InputSignal, +TransformContext, -Output)} will provide a generic view of processing to the user with standarized semantics.

Another complex ``scenario'' is classification. We find this approach suitable to control swiweka effectively. Each classification context can be defined by:

'Classification'(Identifier, ListOfCategories, ListOfDataRecords, Classifier, ClassificationVector)

\begin{itemize}

 \item Identifier gives a label for the classification set
 \item 'Category'(Label, Type, ValuesRange) specifies which type of data is going to be classified
 \item 'DataRecord'(RecordID, ListOfMatchingValues) is a data instance of the dataset 
 \item 'Classifier'(Name, ...)
 \item 'Score'(RecordID, ListOfValues) keeps track of each score (vector of scores depeding of the statistical distribution) for every record.
\end{itemize}

By defining these functors, the engine is able to interpret the data and the user has a clear semantic template to create the classification environment properly. We could even define some constraints and predefined functors instanes for a specific domain.

\subsection{Computation reasoning}\label{subsec:ctl}

The previous abstraction of computations offers a cleaner view of the system without loosing flexibility but there are more issues to be addressed. The system proposed so far does not provide anything novel in respect of other approaches presented in \S 2 and even could seem even more cumbersome to deal with. The major advantage of using a logic framework as basis of the DSP system is the reasoning capability of such a framework.

This reasoning skills are extended to computation reasoning by using the Concurrent Transaction Logic as superset of the Predicates Logic. Transaction Logic defines the logics that allow us to control and manage the execution of the computation and reason about the effects in these terms:

\begin{itemize}
 \item Integration of updating transactions in queries with formal logics
 \item Integration of declarative and procedural meanings of computational transactions (paths)
\end{itemize}

The extended syntax provides connectives for different program states to create a complex execution path (allowing parallel/serial operations, concurrency control, loops, conditionals, recursion and subroutines). The declarative meaning and the procedural meaning are combined:

computation A executes along path X = A is true on X

Actually we are now in position to define several execution paths for the same computation, control the execution of those paths, reason about the results and the truth of such executions and store the computation sequences and data states.

(need one good example here)

\section{SWI-DSP in real applications}\label{sec:realapp}

SWI-DSP defines therefore a logic front-end for computations in DSP ``scenarios''. We have described how Concurrent Transaction Logic converts this framework based on a predicate calculus into a DSP reasoner which control the execution and data of the different scenarios. This system still needs an implementation which allow an external user to control easily the reasoner, access the computations and visualise the results.

\subsection{Languages hosting SWI-Prolog}

Even when there is not current specific work on wrapping SWI-DSP in any other programming language which could make easier to develop a Graphical User Interface (as it probably would restrict the logic capabilities of the framework), we have to mention that SWI-DSP can be wrapped by other languages through the interfaces used for the communication layers (they define both ways of interaction).

@@

Although SWI-Prolog is a well-known and used implementation of Prolog it does not represent the most suitable framework to deal with multimedia data. The multimedia community in general needs a standard machine-readable format with proper semantics to manage data. The Resource Description Framework (RDF) has been proposed as such a format and defines an Standard to represent linked data by means of triples in the form of (Subject, Predicate, Object).

This standard can be implemented with different syntaxes. We prefer to use Notation3 [x] whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs [g]. This is a suitable framework to represent data that can be shared and published in the so-called ``web of data``.

The Ontology Web Language (OWL) provides the framework to express the knowledge of a particular domain in RDF through concepts and relationships. Thus, descripting semantic models related to audio analysis are not restricted to a local implementation but can be shared and processed by different systems.

The project called Henry implements an N3/Prolog parser and reasoner based on Concurrent Transaction Logic. This agent defines an entailment module capable of reasoning over the SWI-DSP computations, which are identified through Universal Resource Identifiers (URIs), and N3 statements and rules. The semantic web package module for SWI-Prolog interprets SPARQL queries which are passed to Henry which produces RDF descriptions of computations by performing the entailment and triggering SWI-DSP computations.

\subsection{Visualisation}\label{subsec:vis}

Audio data can just be poorly interpreted in a float/binary format. Time and frequency domain signals and features in other domains are meaningful for an expert when they are visually represented. Although a logic domain does not seem to be the perfect environment to visualise audio data, we can easily save this difficulty.

By controlling SWI-DSP by means of Henry we can easily convert our Prolog types and binary data into RDF (machine-readable) descriptions where each source of data is represented through a link or as a ''literal`` in a basic datatype format. RDF descriptions can be easily browsed with the new extensions of classical web browsers for linked data. Even more interesting, dedicated agents with an SPARQL point are able to query DSP through the semantic web resources as explained in \ref{subsec:henry}. Thus, we are able to visualise our own computations but also a scalable web of data with shared semantics which is enriched with computations from other local machines.

(I dont know if i should mention SV and the new extensions here....)

@@

We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge. The system design ''breaks apart`` traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner. This reasoner can be extended and enriched by the open source APIs working on the foundations of the system and controlled by ''Prolog-speaking`` agents. SWI-DSP is fairly well positioned for the new web technologies and actually it finds in them the perfect semantic and storing support to overwhelm traditional and closed DSP systems.

@@

@@

>>>>

This layout allow us to wrap fast and efficient implementations of algorithms (in some imperative language) to provide predicates that determine the relationship between defined types (as black box computation) giving a ``functional programming'' view of the system preserving computational efficiency. This functional programming approach (which allow us to create workflows as rules with a procedural meaning) is motivated by the advantages of functional languages like Haskell \cite{haskell} that present a semantically oriented view of computations instead of a machine-oriented one. There are also several works that develop such a typed-functional language on a logic framework like lambda-Prolog or Mercury \cite{mercury}. However, our approach results to be more suitable to develop an open platform for research and development. External engines based on open source APIs provide faster computations than any functional or logic/strong-typed language and a suitable developers interface. Furthermore, SWI-Prolog features a very powerful, interactive and flexible front-end which is also a requirement for the proposed system.

@@

\begin{itemize}
 \item signal: this type represents an audio signal in a physical timeline. 
 \item feature: this type represents some side of the audio signal properties
 \item parameter: this type represents the different variables of the computation context (computations are not only a one-to-one function)
 \item timestamp: this type represent the reference of a data object to a timeline
\end{itemize}

@@

\begin{figure}
\centerline{\framebox{\includegraphics[width=\columnwidth]{workflow.png}}}
\caption{Example of how to create a workflow: compare the tempo of a signal and a reverbered version of it}
\label{fig:workflow}
\end{figure}

Figure \ref{fig:workflow} shows a possible workflow that is designed by using the functional framework formed of datatypes, a standard format for data and the modules of SWI-Prolog predicates which are described in \ref{sec:modules}.

@@

\section{DSP front-end}\label{sec:frontend}

So far we have presented how SWI-DSP provides a vocabulary of logic predicates and datatypes to perform audio analysis imitating a functional language, but this aproach is still inappropiate for such tasks. We have organized the modules in communication layers and task-oriented modules, but there is an important trade-off between understandability and specification. The communication layer even offering an application-independent access to the computation libraries still requires a high knowledge of such libraries besides a SWI-Prolog familiarization which is not expected in most part of the potential users. The datatypes are not the only necessary semantic resource to provide an actual workbench for DSP. We require a front-end that reflects the experimental domain providing an interactive ``laboratory''.

We define some ``DSP scenarios'' that reflect the researching activity. This scenarios define the necessary patterns and datatypes (implemented as predicates and functors) to allow the user to actually perform high-level research preserving the necessary flexibility. The elemental scenarios are:

\begin{itemize}
 \item Transform: This scenario provides a pattern to apply any computation over an audio signal. It specifices by a datatype the framing parameters, the running engine, the engine configuration and parameters, the sample rate (if variable), the type of transform and the domain specifications. These factors are captured with a functor used to link input/output (with full flexibility) in transform(+Input, +Transform, -Output)
 \item Classification: This scenario is provided for a proper setup of classification. This scenario specifies the classification domain, the categories (names, types and ranges), the classification schema and the training set. It also provides a pattern for the input data set which is interpreted along with the classification environment to return the resulting 
\end{itemize}

The patterns and functors are interpreted by hardcoded prior knowledge which starts up the subsequent steps to provide the queried output. More complex scenarios can be created in order to achieve higher or broader semantics for research activity. This interactive logical framework can be interfaced, tranlasted or embedded into other environments by means of the wide range of SWI-Prolog tools (see \ref{subsec:henry}).

\section{Data management and computation reasoning}\label{subsec:datamanage}

We still need to address, probably, the most important issue to accomplish the proposed architecture. We could say that such a system is indeed aimed at managing complex and large data, which is related through computational ``blocks'', and allowing reasoning about it. This goal requires the following achievements: an efficient storage of data objects and the relationships (predicates) linking them, mechanisms to control the concurrency, consistency and side-effects of the computations and a high-level interface to interact with them.

We do not develop such manager within SWI-DSP as the database implementation actually depends on the needs of the host agent interacting with SWI-DSP. However, supposed a database we can describe the architecture in detail. Data is stored in the database by ``tabling'' predicates (which link the input, the context and the output). Tabled predicates (we table predicates that comprise a big load of computations) are stored in the database the first time they are called so data can be retrieved automatically without running the external engines simulating not functional but semantic relationships of complex (audio) data. For this purpose, the data identifiers mechanism seems to be ideal as data ids are linked (and binary data is dumped and loaded from external files relieving memory) and stored wraping complex data as a declarative-oriented units of knowledge. An example of a tabled predicate is:

\begin{verbatim}
?-table(decode('../myfile', signal(2, 44100, 1200000, ['__data_345', '__data_346']))).
\end{verbatim}

The system that was relying in a functional view of computations can be seen now as a knowledge source that can be managed by combining procedural declarations of execution paths and declarative transactions and updates of data by means of Concurrent Transaction Logic. By adopting a formalized mechanism to manage data we are not only achieving a higher level of abstraction but we provide the system with tools to store/export/import data. The system is then open to interact with other environments and applications (saving data format/type conversion issues) allowing a researcher or user to extend his workspace and data resources.

\subsection{Concurrent Transaction Logic}\label{sec:ctl}

Concurrent Transaction Logic \cite{ctl} extends the predicate logic to unify in a declarative formalism the control of an underlying database (SWI-Prolog and other prolog implementations just rely on assert and retract predicates for data updating, being very limited). Concurrent Transaction Logic allows a declarative definition of execution paths and data updates by providing new connectives to express complex data transactions: parallel and serial conjunctions, execution in isolation and executability checking. Rules define possible execution paths for a goal with a higher-level meaning than just declared procedures as we are certainly dealing with actions (and their effects) instead of sequences of predicates.

Therefore, we can prevent side-effects before modifying the data state, reason about the actions and their effects, control the execution of different processes and compare different paths for the same goal. Concurrent transaction logic applied to complex data/computational environments allows and helps to develop ``virtual laboratory workbenches'' \cite{virtual}.

SWI-DSP combined with a data manager based on Concurrent Transaction Logic (CTL) develops a multimedia knowledge system to drive research with semantics and logic formalism exploiting reasoning and declarative definition of computations and providing an interactive and shareable source of knowledge.

@@

We present a related project so-called Henry which develops a real CTL-based agent aimed at creating a distributed multimedia knowledge workspace built in semantic web technologies. The multimedia community needs a standard machine-readable format with proper semantics to manage complex data involved in any sub-domain, so data can be exchanged and shared by different systems enriching the common knowledge and resources.

The Resource Description Framework (RDF) developed as the standard format for new (data-oriented) web technologies expresses linked data by means of triples in the form of (Subject, Predicate, Object). This standard can be implemented with different syntaxes. We prefer Notation3 \cite{n3} whose syntax allow us to write RDF statements in a human-readable way and also to write ``rules'' as named graphs. The Ontology Web Language (OWL) provides the framework to express the knowledge of a particular domain in RDF through concepts and relationships. Thus, descripting semantic models related to audio analysis are not restricted to a local implementation but can be shared and processed by different systems. These elements form a suitable framework to represent and manage data and algorithms in a ``web of (complex) data''.

Henry develops an N3/Prolog parser and reasoner based on CTL. (Text for Yves). Talk about results and performance...
(talk about data visualisation?)

@@

(may have to reword this after all the changes)
We have presented a powerful reasoner for multimedia analysis implemented as a logic-based framework which allow us to convert low-level computations into shareable and distributed knowledge. The system design ''breaks apart`` traditional applications architecture to introduce logic and reasoning machinery on top of the computation engines providing a more intelligent, capable and understandable DSP reasoner. This reasoner can be extended and enriched by the open source APIs working on the foundations of the system and controlled by ''Prolog-speaking`` agents. SWI-DSP is fairly well positioned for the new web technologies and actually it finds in them the perfect semantic and storing support to overwhelm traditional and closed DSP systems.

@@

\bibitem{virtual} Bonner, A.J., Shrufi, A., Rozen, S.: LabFlow-1: a Database Benchmark for High-Throughput Workflow Management. In: 5th International Conference on Extending Database Technology, pp. 463-478. Springer-Verlag, Lecture Notes in Computer Science, volume 1057. Avignon (1996).

@@

\bibitem{mercury} Mercury Home Project \url{http://www.cs.mu.oz.au/research/mercury/index.html}

\bibitem{haskell} Haskell \url{http://www.haskell.org/}

\bibitem{swi} SWI-Prolog Home \url{http://www.swi-prolog.org/}

\bibitem{n3} Notation3 \url{http://www.w3.org/DesignIssues/Notation3}
